Researchers'' perceptions of malaria eradication: findings from a mixed-methods

### 2018-04-10 paper.Rmd
major update

---
title: "Researchers' perceptions of malaria eradication: findings from a mixed-methods analysis of a large online survey"
bibliography: library.bib
output: 
  pdf_document:
    keep_tex: true
    latex_engine: xelatex
    includes:
      in_header: preamble.sty
csl: elsevier-with-titles.csl
---

@@

Quantifying an event's probability and time frame is essential for calculating its expected value. In the case of global malaria eradication, uncertainty regarding feasibility makes it difficult for policymakers and public health practitioners to forecast the expected value of future benefits. The opportunity cost of investments in eradication-specific interventions can be high, particularly in contexts with other urgent health priorities. In a large online survey of malaria researchers (n = 844), we query perceptions regarding the likelihood and time frame of eradication. We assess pessimism/optimism (via the proxy of years-to-eradication), broken down by area of expertise, and adjust for selection bias. We also carry out a qualitative analysis of free-form comments provided by 61\% of respondents. Our results show that researchers are skeptical of the feasibility of near-term malaria eradication, with most believing that eradication will not be achieved for 50 years or more.
\end{changemargin}
\end{center}

@@

Malaria is a parasitic disease transmitted among humans by mosquitoes. The _Plasmodium falciparum_ species accounts for a large majority of the 200 million annual cases as well as the half million annual deaths worldwide [@White2014] [@World2016]. Malaria "elimination", the "interruption of all local transmission of the infection in a country or region" [@Tanner2015] is actively being pursued by dozens of countries around the world, leading to a renewed push for "eradication" ("the worldwide interruption of transmission") [@Hopkins2013] [@Rabinovich_2017] [@TheLancet2011].

This is not the first time that eradication has been in the international spotlight. The scientific and public health communities have had eradication on their policy agenda since the WHO established the Global Malaria Eradication Program in the 1950s [@Njera2011]. In 1957, U.S. President Dwight Eisenhower told Congress that malaria could be expected to be eradicated in five years. Following the failure of the WHO's first attempt, the focus shifted away from global eradication towards local elimination and control strategies.  In recent years, much of the discourse regarding malaria has shifted back to global eradication [@Roberts2007], with funders, researchers, and public health practitioners rallying to the cause [@Tanner2015]. The Bill and Melinda Gates Foundation has begun actively promoting eradication as feasible "within a generation" [@Gates2014], and the World Health Organization (WHO) has set ambitious goals, stating the objective of eliminating malaria in 65 new endemic countries from 2015 through 2030 (Figure 1).

@@

Most of the current research on expert opinion regarding the feasibility of malaria eradication focuses on the \emph{how} rather than the \emph{if} and the \emph{when} [@Tanner2015]. The participants in the Malaria Eradication Research Agenda (MalERA) process, in particular, have guided research goals and identified gaps in order for elmination to occur [@Alonso2011] [@Rabinovich_2017]. Though the MalERA authors firmly state that eradication is \emph{not} feasible given the "current tools and state of knowledge", mentions to the time frame are vague ("within the lifetime of young scientists just embarking on their careers" [@Alonso2011] and "in the longer term" [@Rabinovich_2017]) and no mention is made of the perception of the overall probability of achieving eradication.  International programmes, such as the WHO Global Malaria Programme (GMP), have acknowledged the need "to take an official position on how and under what timeline malaria eradication could be achieved" [@WHO2015]. Clarity on timeline and likelihood of eradication could inform policy and forecasting models, and plays a crucial role in the economic analysis of the expected value of malaria elimination initiatives. But achieving clarity is difficult, given the many complex and interacting variables which affect malaria transmission, research funding, and technological development. The question of \emph{how} is complex enough, rendering questions of \emph{if} and \emph{when} even more difficult to answer.

The general objective of eradication serves to inspire, rally funder support, motivate researchers, and focus the efforts of public health practitioners. Proponents of disease eradication point to the success of historical and current campaigns (smallpox and polio, respectively), and highlight the long term benefits in health and wealth to future generation. One often cited figure for smallpox was a cost-benefit ratio of more than 400 to 1 [@Barrett_2006]; a more recent analysis on River Blindness suggests that the costs of elimination (relative to staying in "control mode") in Africa would be far lower even in the short term, thanks to the improvements it would lead to in both treatment times and prevented surveillance costs [@Kim2017].   However, the opportunity cost of investments in eradication-specific interventions is likely to be high, particularly in contexts of very low resources and many competing health problems. Additionally the estimated impact of malaria on economic indicators is heterogeneous, with singificant impact in areas of Sub-Saharan Africa - where the endemicity is high - and little to no impact in areas of low endemicity [@Cutler_2010] [@Barofsky2015]. Additionally, the “expected value” of an investment in a binary scenario (eradication or not) is a function of the probability of the scenario’s occurrence, the temporal lag of that occurrence, and the uncertainy around both probability of occurrence and temporal lag. This uncertainty makes the "risk" of investing in malaria elimination and eradication high. Therefore, estimating the likelihood and time frame of eradication of malaria is essential for making sound investments in health. But how can we prospectively quantify these values given the great deal of uncertainty surrounding eradication?

Time to eradication, a crucial component in any hypothetical cost-benefit analysis that takes into account discounting, is difficult to assess and requires several layers of assumptions in order to generate even a range of possible scenarios; this can be done with a high degree of reliability for diseases closer to eradication such as Lympathic Filariasis [@Kastner2015], but the uncertainties regarding malaria's future treatment protocols and epidemiology are too great to allow for a robust analysis based on cascading assumptions. Many studies have shown value in expert elicitation as a means to reduce uncertainty and inform decision making [@Morgan_2014]. But experts can sometimes be mistaken, a fact implicitly recognized when patients request a "second opinion" from another doctor when facing uncertainty. Additionally, relying on only a small panel of experts for a problem as complex as malaria exposes one to bias, just as an overly concentrated investment portfolio exposes one to more risk. As Sir Francis Galton demonstrated in his famous study in which he showed that the crowds' aggregated estimates of cow's weight formed a quasi-normal distribution centered around the true weight [@Wallis2014] [@Galton1907Voxpopuli], averaging the perceptions of many can be more accurate than taking the opinion of any single expert, since the biases of diverse viewpoints can be complementary and symbiotic.

Assessing the likelihood and time frame of eradication, therefore, requires the combination of multiple points of view. Measuring consensus and discord among disease-specific researchers from a variety of disciplines can serve as a barometer of (informed) opinion, both guiding resources and identifying areas of concern [@Keenan2013]. We propose that the aggregation of malaria researchers' perceptions regarding the time frame andlikelihood of eradication forms a probability distribution which can be used to estimate the expected value of eradication-specific investments.

A great deal of previous research from serveral endemic areas already covers the cost per case prevented [@Sicuri2011] [@Silumbe2015] [@Btto2016] [@IlungaIlunga2014] [@Dalaba2015]. Likewise, literature exists which could serve as a model for quantifying the location-specific opportunity costs associated with funneling funds towards malaria eradication [@Stuckey2014] [@White2011] [@Korenromp2012]. The correct discount rate for estimating the value of future lives saved is more of a philosophical question than an economic one. This leaves only the probability and time-frame to eradication, questions which have been addressed anecdotally, but never answered quantitatively.

Though the case-specific marginal cost of eradication can be expected to be high (relative to a simple control approach), successful eradication would mean massive recurring savings in the long-term [@Barrett2013]. However, to the extent that the case-specific marginal cost of prevention in an eradication campaign is high, estimating the likelihood of success is fundamental to the correct distribution of resources, particularly in low-income environments.

In other words, the rational assignment of resources for malaria eradication campaigns hinges on the expected value of those campaigns, the latter being a direct function of the discount rate applied to future benefits (which this study does not aspire to assess) and the probability of "success" (ie, eradication). Holding constant factors such as the cost of eradication and the benefits of achieving it, the return on investment is a linear function of eradication's probability. Given this, estimating probability is crucial to evaluating if and when attempts at eradication should be undertaken.

The objective of this study is to gauge (expert) opinion about, estimate the likelihood of, and quantify the potential time frame to malaria eradication through a systematic online survey of malaria research professionals from a wide array of academic disciplines. In doing so, we aim to help guide the optimal distribution of health resources by informing estimations of the expected value of malaria eradication efforts.

@@

The survey was intentionally as short as possible, so as to appeal to time-pressed participants. However, supplementary data on researchers is of value for the assessing of selection bias and determinants of perception. Accordingly, we estimated the likelihood of a participant being of one of two genders (male or female) based on first name. In order to do this, we used data from the North Atlantic Population Project, and U.S. government [@Lincoln]. We also calculated the total number of citations which each author had received, as well as the total number of articles authored. We binned total citations into three categories: 0-5 (junior-level researchers, PhD students, etc.); 6-99 (most professional academics); and >99 (the most prolific researchers). The searching and retrieval of information pertaining to articles and citations from the PubMed database was carried out using the RISmed package [@Kovalchik2015].

@@

Quantitative analysis was carried out in R language [@R]. Ordinary least squares regression was employed to estimate the effect of area of expertise on perceived years to eradication. All analysis code, as well as the code used for the identification and contacting of participants, is publicly available at [https://github.com/joebrew/malaria_survey](https://github.com/joebrew/malaria_survey).

@@

Many commenters thought that the terms "eradication" and "malaria" were so complex and nebulous that the public health practitioners should avoid the them all together, so as to not repeat the mistakes of the WHO GMP in the 1950s. For example, some highlighted that talking about "malaria" as one disease misses the mark, since the different species of parasite and contexts in which they live make elimination in each area very distinct from other areas.  One commenter wrote that eradication is a "postwar" idea that developed from the "abandonment of a broad sociopolitical understanding of the causes of disease, and the emphasis on technological solutions." Many stated that global malaria eradication was simply not possible, and two argued that it may not be desirable or ethical. One stated that the concept was "absurd" and that "I'm not even sure why people talk about it". Another questioned the utility of discussing "eradiction" as a concept:

@@

Approximately three-quarters of respondents believe that malaria will not be eradicated in the next 50 years; in other words, given that our pool of respondents is of typical post-PhD age, most believe they will not live to see eradication. When adjusted for selection bias, the perceived likelihood of eradication 50 years from now remains similarly low, but estimates for shorter-term eradication are even lower.

In regards to differences by academic discipline, it is important to note that diversity within disciplines was greater than diversity between them. That said, one might surmise as to why disciplines at the extremes differed. Chemists, on average, reported that eradication will be achieved in less than 40 years, whereas economists thought it would take a full ten years more.  Speculatively, the collective optimism in the field of chemistry may be attributable to recent advances in drug development and insecticides, whereas economists may be more pessimistic due to concerns such as free-riding when it comes to global public goods [@barrett] [@Bodansky_2012]. This study was not designed to generate inference regarding reasons why disciplines might differ in perception, but it does suggest further lines of research.

@@

This paper has several limitations. First, academic researchers are specialists - their narrow, field-specific view of eradication's feasibility is of arguable reliability, given that they may be unfamiliar with the operational, cultural and "real-world" challenges of malaria eradication. Though crowds have been found to be more "wise" than individuals in many cases, the application of an approach similar to the "wisdom of crowds" is not suitable to all classes of problems [@Mannes_2014]. Crowds can be susceptible to social biases [@Lorenz_2011] (although this survey's anonimity largely protects against this issue), and other biases may come in to play, especially given that our study was of a crowd of "specialists", rather than the population as a whole.

@@

As with Keenan et al's survey of experts regarding the feasibility of eradication of neglected tropical diseases [@Keenan2013], our survey detected relatively high levels of eradication skepticism and did not delve into whether researchers had clinical or operational experience, nor did it assess opinions of program workers, nor did it explore complexities pertaining to different types of or forms of existence of malaria. Though our sample size was over twice Keenan's, this was largely due to having contacted more authors, as our response rate was only one fourth as large.

@@

Our survey attempted to quantify the likelihood of, and time frame to, malaria eradication by gauging the collective opinion of malaria researchers.  The findings of our survey show researchers expressing hesitance about the likelihood of eradication and suggesting a long time frame. The causes for scepticism are diverse, but common themes were the need for innovation, systemic challenges, and the complexity of the disease and its transmission.

The implication of these results are two-fold: (1) that those working or investing in eradication-specific campaigns, as well as those modelling these campaigns' hypothetical cost-effectiveness, should factor in researchers' perceived low likelihood and long time frame when calculating those campaigns' expected value; (2) that champions of near-term eradication may need to make a more compelling case to malaria researchers of eradication's feasibility, in order to better focus and inspire the latter.

>>>>

---
title: 'Researchers'' perceptions of malaria eradication: findings from a mixed-methods
  analysis of a large online survey'
csl: elsevier-with-titles.csl
output:
  pdf_document:
    includes:
      in_header: preamble.sty
    keep_tex: yes
    latex_engine: xelatex
  word_document: default
bibliography: library.bib
---

@@

Quantifying an event's probability and time frame is essential for calculating its expected value. Uuncertainty regarding feasibility makes it difficult for policymakers and public health practitioners to forecast the return on investment of malaria eradication. Since the opportunity cost of investments in eradication-specific interventions can be high - particularly in contexts with other urgent health priorities - understanding eradication's expected value is essential. In a large online survey of malaria researchers (n = 844), we query perceptions regarding the likelihood and time frame of eradication. We assess pessimism/optimism (via the proxy of years-to-eradication), and adjust for selection bias. We also carry out a qualitative analysis of the free-form comments provided by 61\% of respondents. Our results show that researchers are skeptical of the feasibility of near-term malaria eradication, with most believing that eradication will not be achieved for 50 years or more.
\end{changemargin}
\end{center}

@@

Malaria is a parasitic disease transmitted among humans by mosquitoes. The _Plasmodium falciparum_ species accounts for a large majority of the 200 million annual cases as well as the half million annual deaths worldwide [@White2014] [@World2016] [@Ashley2018]. Malaria "elimination", the "interruption of all local transmission of the infection in a country or region" [@Tanner2015] is actively being pursued by dozens of countries around the world, leading to a renewed push for "eradication" ("the worldwide interruption of transmission") [@Hopkins2013] [@Rabinovich_2017] [@TheLancet2011].

This is not the first time that eradication has been in the international spotlight. The scientific and public health communities have had eradication on their policy agenda since the World Health Organisation (WHO) established the Global Malaria Eradication Program in the 1950s [@Njera2011]. In 1957, U.S. President Dwight Eisenhower told Congress that malaria could be expected to be eradicated in five years. Following the failure of the WHO's first attempt, the focus shifted away from global eradication towards local elimination and control strategies.  In recent years, much of the discourse regarding malaria has shifted back to global eradication [@Roberts2007], with funders, researchers, and public health practitioners rallying to the cause [@Tanner2015]. The Bill and Melinda Gates Foundation has begun actively promoting eradication as feasible "within a generation" [@Gates2014], and the WHO has set ambitious goals, stating the objective of eliminating malaria in 65 new endemic countries from 2015 through 2030 (Figure 1).

@@

Most of the current research on expert opinion regarding the feasibility of malaria eradication focuses on the \emph{how} rather than the \emph{if} and the \emph{when} [@Tanner2015]. The participants in the Malaria Eradication Research Agenda (MalERA) process, in particular, have guided research goals and identified gaps in order for elmination to occur [@Alonso2011] [@Rabinovich_2017]. Though the MalERA authors firmly state that eradication is \emph{not} feasible given the "current tools and state of knowledge", mentions to the time frame are vague ("within the lifetime of young scientists just embarking on their careers" [@Alonso2011] and "in the longer term" [@Rabinovich_2017]) and no mention is made of the perception of the overall probability of achieving eradication.  International programmes, such as the WHO Global Malaria Programme (GMP), have acknowledged the need "to take an official position on how and under what timeline malaria eradication could be achieved" [@WHO2015]. Clarity on timeline and likelihood of eradication could inform forecasting of disease transmission, and plays a crucial role in the economic analysis of the expected value of malaria eradication initiatives, ultimately informing health policy decisions. But achieving clarity is difficult, given the many complex and interacting variables which affect malaria transmission, research funding, and technological development. The question of \emph{how} is complex enough, rendering questions of \emph{if} and \emph{when} even more difficult to answer.

Proponents of disease eradication point to the success of historical and current campaigns (smallpox and polio, respectively), and highlight the long term benefits in health and wealth to future generations [@Barrett2013]. One often cited figure is for smallpox eradication, in which the investment of each dollar led to a return of greater than 400 [@Barrett_2006]; unlike smallpox (for which analysis was able to be retrospectively constructed based on observed costs and benefits), in the case of a prevalent disease, forecasting models have to be used in order to prospectively estimate the value of an investment.  \st{Additionally the estimated impact of malaria on economic indicators is heterogeneous, with singificant impact in areas of Sub-Saharan Africa - where the endemicity is high - and little to no impact in areas of low endemicity} [@Cutler_2010] [@Barofsky2015].

The general objective of eradication serves to inspire, rally funder support, motivate researchers, and focus the efforts of public health practitioners. But to the extent that malaria eradication (by definition) has never occurred, the parameters needed for a ex-post cost-benefit analysis are unknown. Nonetheless, prospectively estimating eradication's return on investment is crucial to deciding when and how to pursue the goal, especially in light of the high direct and opportunity costs of eradication-specific interventions. One approach to cost-benefit analysis is the use of mathematical and stochastic models. These have been applied to diseases which are closer to eradication than malaria, since the uncertainty around model input parameters is less, requiring fewer cascading assumptions in order to present possible comparative scenarios. For example, Kastner et al. were able to describe 4 relatively realistic pathways to Lympathic Filariasis eradication, as well as the pre-requisite role and magnitude of certain interventions [@Kastner2015]. A similar modeling framework was then used to estimate the cost of eradication [@Kastner2017]. A recent modeling analysis on onchocerciasis eradication, based on a disease transmission model, showed that the costs of elimination (relative to staying in "control mode") in Africa would be far lower even in the short term, thanks to the improvements it would lead to in both treatment times and prevented surveillance costs. Given the relative proximity of eradication, and the narrow geographic scope, the authors were also able to estimate the timeline to eradication [@Kim2017].

This level of detail and specificity in modeling the cost-benefit landscape of malaria eradication, unfortunately, is not possible (or at least not commonly undertaken), given its high prevalence and epidemiological complexity. In fact, there are no mathematical models (to the authors' knowledge) estimating the likelihood and time frame of eradication, or its derivative cost-benefit ratio. While a great deal of rigorous studies exist forecasting the effect of certain tools and strategies relative to alternatives [@Brady2017] or aimed at regional disease elimination [@Kim2015], these analyses do not endeavor to estimate the generalization of those tools or strategies to other areas. Globally, where mathematical modeling has been used for the purposes of forecasting the future epidemiology of malaria, the methods have generally been aimed at optimizing elimination methods [@Slater2015], determining whether a strategy is scalable [@Brady2017], guiding funding and drug development [@Patouillard2017] [@Slater2017], or comparing a range of hypothetical morbidity scenarios [@Winskill2017], rather than assessing the likelihood of or time until the occurrence of those scenarios. In other words, researchers are (rightfully) focused on the \emph{how} of eradication, rather than the forecasting of its timeframe or likelihood, the lack of attention paid to the latter likely being a reflection of its complexity, not lack of importance.

To the extent that (i) estimating eradication's likelihood and timeframe is essential to forecasting the cost-benefit ratio of eradication interventions, and (ii) mathematical models generally do not venture into areas of such high uncertainty, alternative methods are needed. Just as the stock market aggregates perceptions to provide a fairly accurate assessment of something as complex as a company's value, aggregating perceptions may be a useful tool for tackling the complexity of malaria eradication. Many studies have shown value in expert elicitation as a means to reduce uncertainty and inform decision making [@Morgan_2014]. \st{But experts can sometimes be mistaken, a fact implicitly recognized when patients request a "second opinion" from another doctor when facing uncertainty. Additionally, relying on only a small panel of experts for a problem as complex as malaria exposes one to bias, just as an overly concentrated investment portfolio exposes one to more risk.} As Sir Francis Galton demonstrated in his famous study in which he showed that the crowds' aggregated estimates of cow's weight formed a quasi-normal distribution centered around the true weight [@Wallis2014] [@Galton1907Voxpopuli], averaging the perceptions of many can be more accurate than taking the opinion of any single expert, since the biases of diverse viewpoints can be complementary and symbiotic. Measuring consensus and discord among disease-specific researchers from a variety of disciplines can serve as a barometer of (informed) opinion, both guiding resources and identifying areas of concern [@Keenan2013].

The optimal assignment of resources for malaria eradication campaigns hinges on the expected value of those campaigns, the latter being a direct function of the discount rate applied to future benefits and the probability of "success" (ie, eradication). Holding constant factors such as the cost of eradication and the benefits of achieving it, the return on investment of malaria eradication initiatives is a function of eradication's timeframe (assuming a > 0 discount rate) and probability (assuming a < 100% likelihood of success). Given this, estimating these parameters is crucial to evaluating if and when attempts at eradication should be undertaken.

We propose that the aggregation of malaria researchers' perceptions regarding the time frame and likelihood of eradication forms a probability distribution which can be used to estimate the expected value of eradication-specific investments. The objective of this study is to gauge (expert) opinion about, estimate the likelihood of, and quantify the potential time frame to malaria eradication through a systematic online survey of malaria research professionals from a wide array of academic disciplines. In doing so, we aim to help guide the optimal distribution of health resources by informing estimations of the expected value of malaria eradication efforts.

\st{A great deal of previous research from serveral endemic areas already covers the cost per case prevented [@Sicuri2011] [@Silumbe2015] [@Btto2016] [@IlungaIlunga2014] [@Dalaba2015]. Likewise, literature exists which could serve as a model for quantifying the location-specific opportunity costs associated with funneling funds towards malaria eradication [@Stuckey2014] [@White2011] [@Korenromp2012]. The correct discount rate for estimating the value of future lives saved is more of a philosophical question than an economic one. This leaves only the probability and time-frame to eradication, questions which have been addressed anecdotally, but never answered quantitatively.}

\st{Though the case-specific marginal cost of eradication can be expected to be high (relative to a simple control approach), successful eradication would mean massive recurring savings in the long-term [@Barrett2013]. However, to the extent that the case-specific marginal cost of prevention in an eradication campaign is high, estimating the likelihood of success is fundamental to the correct distribution of resources, particularly in low-income environments. }

@@

The survey was intentionally as short as possible, so as to appeal to time-pressed participants. However, supplementary data on researchers is of value for the assessing of selection bias and determinants of perception. Accordingly, we estimated the likelihood of a participant being of one of two genders (male or female) based on first name. In order to do this, we used data from the North Atlantic Population Project, and U.S. government [@Lincoln]. We also calculated the total number of citations which each author had received, as well as the total number of articles authored. We binned total citations into three categories: 0-5 (junior-level researchers, PhD students, etc.); 6-99 (most professional academics); and >99 (the most prolific researchers). The searching and retrieval of information pertaining to articles and citations from the PubMed database was carried out using the RISmed package [@Kovalchik2015]. Information retrieved about authors was used to de-bias paramaters in an ordinary least squares regression, analyzing the association between number of publications and perceived years to eradication.

@@

Quantitative analysis was carried out in R language [@R]. All analysis code, as well as the code used for the identification and contacting of participants, is publicly available at [https://github.com/joebrew/malaria_survey](https://github.com/joebrew/malaria_survey).

@@

Many commenters thought that the terms "eradication" and "malaria" were so complex and nebulous that the public health practitioners should avoid the them all together, so as to not repeat the mistakes of the WHO GMP in the 1950s. For example, some highlighted that talking about "malaria" as one disease misses the mark, since the different species of parasite and contexts in which they live make elimination in each area very distinct from other areas.  One commenter wrote that eradication is a "postwar" idea that developed from the "abandonment of a broad sociopolitical understanding of the causes of disease, and the emphasis on technological solutions." Many stated that global malaria eradication was simply not possible, and two argued that it may not be desirable or ethical. One stated that the concept was "absurd" and that "I'm not even sure why people talk about it". Another questioned the utility of discussing "eradication" as a concept:

@@

Approximately three-quarters of respondents believe that malaria will not be eradicated in the next 50 years. When adjusted for selection bias, the perceived likelihood of eradication in 50 years remains similarly low, but estimates for shorter-term eradication are even lower.

@@

This paper has several limitations. Conceptually, academic researchers are specialists - their narrow, field-specific view of eradication's feasibility is of arguable reliability, given that they may be unfamiliar with the operational, cultural and "real-world" challenges of malaria eradication. Though crowds have been found to be more "wise" than individuals in many cases, the application of an approach similar to the "wisdom of crowds" is not suitable to all classes of problems [@Mannes_2014]. Crowds can be susceptible to social biases [@Lorenz_2011] (although this survey's anonimity largely protects against this issue), and other biases may come in to play, especially given that our study was of a crowd of "specialists", rather than the population as a whole.

@@

As with Keenan et al's survey of experts regarding the feasibility of eradication of neglected tropical diseases, our survey detected relatively high levels of eradication skepticism and did not delve into whether researchers had clinical or operational experience, nor did it assess opinions of program workers, nor did it explore complexities pertaining to different types of or forms of existence of malaria. Though our sample size was over twice Keenan's, this was largely due to having contacted more authors, as our response rate was only one fourth as large [@Keenan2013].

@@

Our survey attempted to quantify the likelihood of, and time frame to, malaria eradication by gauging the collective opinion of malaria researchers.  The findings of our survey show researchers expressing hesitance about the likelihood of eradication and suggesting a long time frame until it is achieved. The causes for scepticism are diverse, but common themes were the need for innovation, systemic challenges, and the complexity of the disease and its transmission.

The implication of these results are two-fold: (1) that those working or investing in eradication-specific campaigns, as well as those modeling these campaigns' hypothetical cost-effectiveness, should factor in researchers' perceived low likelihood and long time frame when calculating those campaigns' expected value; (2) that champions of near-term eradication may need to make a more compelling case to malaria researchers of eradication's feasibility, in order to better focus and inspire the latter.
